---
title: "Comparison Loss"
author: "Cora Allen-Coleman"
date: "2/2/2018"
output: pdf_document
---
#Comparison Loss
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
raw_data <- read.csv("/Users/cora/Dropbox/UW-Madison/CurrentResearch/Ranking/presentations/rankings_introduction/LBW.csv", header = TRUE)
m1_est <- read.csv("/Users/cora/Dropbox/UW-Madison/CurrentResearch/Ranking/presentations/rankings_introduction/m1_estimates.csv", header = TRUE)
require(rstan)
require(rethinking)
```

based on work by P. I. Jewett, L. Zhu, B. Huang, E. Feuer, R. E. Gangnon. “Optimal Point Estimates and Credible Intervals for Ranking County Health Indices.” (under review)

##Create Model with Intercepts for Each County
```{r random intercepts modelmodel}
m1 <- map2stan(
  alist(
    NumLBW ~ dbinom(NumBirths, p) ,
    logit(p) <- a_county[County] ,
    a_county[County] ~ dnorm(a, sigma) ,
    a ~ dnorm(0, 10) ,
    sigma ~ dcauchy(0,1)
  ), data=raw_data, iter=4000, chains=4, WAIC=FALSE) #remove WAIC to speed up
```

##Summarize Posterior
```{r samples from posterior}
## Get Samples From the Posterior for Each County ##
samples = 10000
post <- extract.samples(m1)
post$a_county #each of these is logit(p_i)
ranks <- apply(post$a_county, 1, rank)


## Posterior probabilities of being assigned every rank for each item ##
#summaries of that posterior distribution
apply(ranks,1,quantile)
#mean ranks for each county
apply(ranks,1,mean)
#optimal ranks using squared error loss
rank(apply(ranks,1,mean))
```

##Position Loss Matrix
```{r squared error positional loss and optimization}
# Squared Error Loss Rank
# mean((ranks[1,]-1)^2) #for county 1 ranked at position 1
# mean((ranks[1,]-3)^2)#for county 1 ranked at position 3
SqErrPositionalLoss <- matrix(NA,21,21)
for (i in 1:21) {
  for (j in 1:21) {
    SqErrPositionalLoss[i,j] <- mean((ranks[i,]-j)^2)
  }
}
## b. Find optimal ranking
#using the solve_LSAP function in the clue package
library(clue)
solve_LSAP(SqErrPositionalLoss)
```

##Creating Comparison Loss Matrix
loss associated with placing unit i before unit j
```{r absolute value comparison loss}
AbsComparisonLoss <- matrix(NA,21,21)
for (i in 1:21) {
  for (j in 1:21) {
    #AbsComparisonLoss[i,j] <- abs(prob(rank(i) > rank(j)))
    AbsComparisonLoss[i,j] <- mean(abs(ranks[i,] - ranks[j,]))
  }
}

#zero one mean(rank[i,]>rank[j,])
```

loss associated with placing unit i before unit j
```{r squared error comparison loss}
SqErrorComparisonLoss <- matrix(NA,21,21)
for (i in 1:21) {
  for (j in 1:21) {
    SqErrorComparisonLoss[i,j] <- mean((ranks[i,]-ranks[j,])^2)
  }
}
```

##Comparison Optimization
"local search based on pairwise exchanges, a simple heuristic algorithm based on pairwise swapping of elements to find a local optimum" (see references) TODO
```{r optimization}
#begin optimization with positional ranking

#for each i in 1:21
  #check if pairwise swapping would improve rank

```

#Other Optimization Strategies
```{r}

```
